---
title: "Template"
<<<<<<< HEAD
<<<<<<< HEAD
<<<<<<< HEAD
author: "Omar Zakkouralashy (2889420), Anna Mishmish-Olaru (2881939), Abhigyan Hazarika (student number), Andrea Tzima or Xhima,Raluca Naumcef(2888853), kyle"
=======
<<<<<<< HEAD
author: "Omar Zakkouralashy (2889420), Anna Olaru (2881939), Abhigyan Hazarika (student number), Andrea Tzima or Xhima,Raluca Naumcef, kyle"
=======
author: "Omar Zakkouralashy (2889420), Anna Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima,Raluca Naumcef, Kyle van Emmerik (2897218)"
>>>>>>> b62ff297222185e7bc4e016fd486e3fee4b2a663
=======
=======
author: "Omar Zakkouralashy (2889420), Anna Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima,Raluca Naumcef, Kyle van Emmerik (2897218)"
=======
>>>>>>> b62ff297222185e7bc4e016fd486e3fee4b2a663
author: "Omar Zakkouralashy (2889420), Anna Mishmish-Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima (2885530),Raluca Naumcef, Kyle van Emmerik (2897218)"
>>>>>>> d0097343dd07037992a3f4c54a811f9a9756b238
>>>>>>> a66092d3728bb9fa977f156a6f3c809c2eb7c25b
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
csl: apa.csl

---

# Set-up your environment

```{r package_install, include=FALSE}
install.packages("tidyverse")
install.packages("yaml")
install.packages("rmarkdown")
install.packages("maps")
install.packages("eurostat")
```

```{r packages}
require(tidyverse)
require(rmarkdown)
require(yaml)
library(tidyr)
library(writexl)
library(readxl)
require()
```

# Title Page
The effect of COVID on household savings in the European Union - A social problem.

Omar, Anna, Abhigyan, Raluca, Andrea, Kyle

TG 5 - G 1

Tutorial teacher: Simon 

Simon Loop
Simon Loop

# Part 1 - Identify a Social Problem



## 1.1 Describe the Social Problem

Include the following:

-   Why is this relevant? [@smith2020]

-   ...
# Part 2 - Data Sourcing

## 2.1 Load in the data

```{r loading_data}
library(maps)
library(eurostat)
library(grid)

dat_savings   <- get_eurostat("nasa_10_ki",    time_format = "num", cache = TRUE)
dat_inflation <- get_eurostat("prc_hicp_ainr", time_format = "num", cache = TRUE)
```

## 2.2 Provide a short summary of the dataset(s)

```{r summary}
dataset_summary <- data.frame(
  Dataset    = c("nasa_10_ki", "prc_hicp_ainr"),
  Source     = c("Eurostat", "Eurostat"),
  Indicator  = c("Gross household saving rate",
                 "HICP inflation rate"),
  Unit       = c("% of gross disposable income",
                 "Annual average % change"),
  Level      = c("Country-year", "Country-year"),
  Years_used = c("2016-2022", "2016-2022"),
  Limitation = c("Country-level only, no household-level detail",
                 "National average, not household-specific")
)

dataset_summary
```

#write here about metadata: both datasets come from Eurostat; saving rate is sourced from national accounts (ESA 2010) and measures gross household saving as a share of disposable income; inflation comes from HICP price-index data and measures annual average price change across consumer goods and services xyz

## 2.3 Describe the type of variables included

#write here: saving rate and inflation are ratio-level numeric variables; country is nominal; year is interval; covid_phase and region_group are ordinal/categorical; saving rate data comes from administrative national accounts; inflation comes from HICP price index data, not surveys xyz

---

# Part 3 - Quantifying

## 3.1 Data cleaning

```{r data_cleaning_savings}
eu27 <- c(
  "AT", "BE", "BG", "HR", "CY", "CZ", "DK", "EE", "FI", "FR",
  "DE", "EL", "HU", "IE", "IT", "LV", "LT", "LU", "MT", "NL",
  "PL", "PT", "RO", "SK", "SI", "ES", "SE"
)

years_used <- 2016:2022

savings_clean <- dat_savings %>%
  filter(
    na_item == "SRG_S14_S15",
    sector  == "S14_S15",
    unit    == "PC",
    geo     %in% eu27
  ) %>%
  mutate(year = as.integer(TIME_PERIOD)) %>%
  filter(year %in% years_used) %>%
  select(geo, year, saving_rate = values) %>%
  drop_na(saving_rate)
```

```{r data_cleaning_inflation}
inflation_clean <- dat_inflation %>%
  filter(
    coicop18 == "AP",
    unit     == "RCH_A_AVG",
    geo      %in% eu27
  ) %>%
  mutate(year = as.integer(TIME_PERIOD)) %>%
  filter(year %in% years_used) %>%
  select(geo, year, inflation_rate = values) %>%
  drop_na(inflation_rate)
```

```{r data_cleaning_merge}
combined <- left_join(savings_clean, inflation_clean, by = c("geo", "year"))
```

```{r data_cleaning_check}
combined %>%
  summarise(
    countries         = n_distinct(geo),
    first_year        = min(year),
    last_year         = max(year),
    observations      = n(),
    missing_saving    = sum(is.na(saving_rate)),
    missing_inflation = sum(is.na(inflation_rate))
  )
```

## 3.2 Generate necessary variables

```{r gen_var1}
country_names <- c(
  AT = "Austria",     BE = "Belgium",      BG = "Bulgaria",
  HR = "Croatia",     CY = "Cyprus",       CZ = "Czech Republic",
  DK = "Denmark",     EE = "Estonia",      FI = "Finland",
  FR = "France",      DE = "Germany",      EL = "Greece",
  HU = "Hungary",     IE = "Ireland",      IT = "Italy",
  LV = "Latvia",      LT = "Lithuania",    LU = "Luxembourg",
  MT = "Malta",       NL = "Netherlands",  PL = "Poland",
  PT = "Portugal",    RO = "Romania",      SK = "Slovakia",
  SI = "Slovenia",    ES = "Spain",        SE = "Sweden"
)

combined <- combined %>%
  mutate(
    region = country_names[as.character(geo)],
    region_group = case_when(
      geo %in% c("AT", "BE", "DE", "FR", "IE", "LU", "NL") ~ "Western Europe",
      geo %in% c("DK", "FI", "SE")                          ~ "Northern Europe",
      geo %in% c("CY", "EL", "ES", "IT", "MT", "PT")        ~ "Southern Europe",
      geo %in% c("BG", "CZ", "EE", "HR", "HU", "LV",
                 "LT", "PL", "RO", "SI", "SK")               ~ "Central/Eastern Europe"
    ),
    covid_phase = case_when(
      year <= 2019 ~ "Pre-COVID",
      year == 2020 ~ "COVID shock",
      year >= 2021 ~ "Recovery / inflation period"
    ),
    covid_phase = factor(covid_phase, levels = c(
      "Pre-COVID",
      "COVID shock",
      "Recovery / inflation period"
    )),
    map_period = case_when(
      year == 2019 ~ "Pre-COVID (2019)",
      year == 2020 ~ "COVID shock (2020)",
      year == 2021 ~ "Post-COVID (2021)",
      year == 2022 ~ "Post-COVID (2022)"
    ),
    map_period = factor(map_period, levels = c(
      "Pre-COVID (2019)",
      "COVID shock (2020)",
      "Post-COVID (2021)",
      "Post-COVID (2022)"
    ))
  ) %>%
  drop_na(region, region_group)
```

```{r gen_var2}
baseline_2019 <- combined %>%
  filter(year == 2019) %>%
  select(geo, saving_rate_2019 = saving_rate)

combined <- combined %>%
  left_join(baseline_2019, by = "geo") %>%
  mutate(change_from_2019 = saving_rate - saving_rate_2019)

savings_clean <- combined
```

## 3.3 Visualize temporal variation

```{r temporal_plot}
eu_mean <- combined %>%
  group_by(year) %>%
  summarise(
    mean_saving_rate    = mean(saving_rate,    na.rm = TRUE),
    mean_inflation_rate = mean(inflation_rate, na.rm = TRUE),
    n_countries         = n_distinct(geo),
    .groups = "drop"
  )

ggplot(eu_mean, aes(x = year, y = mean_saving_rate)) +
  geom_line(color = "#2C5F8A", linewidth = 1) +
  geom_point(color = "#2C5F8A", size = 3) +
  annotate("rect", xmin = 2019.5, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.08) +
  geom_vline(xintercept = 2020, linetype = "dashed",
             color = "red", linewidth = 0.7) +
  annotate("text", x = 2020.1, y = 10,
           label = "COVID-19 (2020)", color = "red",
           size = 3.2, hjust = 0) +
  labs(
    title    = "EU27 Mean Household Saving Rate (2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "Year",
    y        = "Mean saving rate (%)",
    caption  = "Note: Unweighted mean across EU27 countries with available data."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    plot.caption  = element_text(color = "grey50", size = 8, hjust = 0)
  ) +
  scale_x_continuous(breaks = 2016:2022)
```

#write here about temporal variation: the EU27 mean saving rate spiked sharply in 2020 as COVID lockdowns suppressed consumption, then gradually declined as economies reopened and inflation eroded disposable income xyz

## 3.4 Visualize spatial variation

```{r visualise_map, fig.width=14, fig.height=10}
europe_map <- map_data("world")

unmatched <- combined %>%
  distinct(region) %>%
  anti_join(europe_map %>% distinct(region), by = "region")
unmatched

map_joined <- left_join(europe_map, combined, by = "region") %>%
  filter(!is.na(map_period))

ggplot(map_joined, aes(x = long, y = lat, group = group, fill = saving_rate)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(ratio = 1.3, xlim = c(-25, 45), ylim = c(34, 72)) +
  scale_fill_gradient2(
    low      = "darkred",
    mid      = "yellow",
    high     = "darkgreen",
    midpoint = median(combined$saving_rate, na.rm = TRUE),
    na.value = "grey90",
    name     = "Saving rate (%)"
  ) +
  facet_wrap(~ map_period, ncol = 2, drop = TRUE) +
  labs(
    title    = "Gross Household Saving Rate across EU27 (2019-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki) | Grey = no data available",
    x = NULL, y = NULL,
    caption  = "Saving rate = gross household saving as % of disposable income. Midpoint set at EU27 median."
  ) +
  theme_void() +
  theme(
    plot.title       = element_text(face = "bold", size = 14),
    plot.subtitle    = element_text(color = "grey40", size = 11),
    plot.caption     = element_text(color = "grey50", size = 8, hjust = 0),
    strip.text       = element_text(face = "bold", size = 12),
    legend.position  = "bottom",
    legend.key.width = grid::unit(2, "cm")
  )
```

#write here about spatial variation: in 2020 the map turns greener across all countries as lockdowns forced saving; by 2022 central and eastern countries turn red as inflation eroded household income xyz

## 3.5 Visualize sub-population variation

```{r visualise_subpopulations}
combined %>%
  ggplot(aes(x = covid_phase, y = saving_rate, fill = region_group)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title    = "Saving Rate by Region Group and COVID Phase (EU27, 2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "COVID phase",
    y        = "Gross saving rate (%)",
    fill     = "Region group",
    caption  = "Groups are broad geographical categories and do not imply causal explanations."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(color = "grey40", size = 10),
    plot.caption    = element_text(color = "grey50", size = 8, hjust = 0),
    axis.text.x     = element_text(angle = 20, hjust = 1),
    legend.position = "bottom"
  )
```

#write here about sub-population variation: western and northern Europe consistently save more than southern and central/eastern Europe; the COVID shock boosted saving across all groups in 2020 but recovery has been uneven xyz hazafreaky

## 3.6 Event analysis

```{r analysis}
combined %>%
  ggplot(aes(x = covid_phase, y = saving_rate, fill = covid_phase)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(
    values = c(
      "Pre-COVID"                   = "#2C5F8A",
      "COVID shock"                 = "#C0392B",
      "Recovery / inflation period" = "#E07B39"
    ),
    guide = "none"
  ) +
  labs(
    title    = "Household Saving Rate around the COVID-19 Shock (EU27, 2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "Period",
    y        = "Gross saving rate (%)",
    caption  = "This plot shows association only; it does not prove COVID caused the observed changes."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    plot.caption  = element_text(color = "grey50", size = 8, hjust = 0)
  )
```

```{r change_from_2019_plot}
combined %>%
  filter(year == 2020) %>%
  ggplot(aes(x = reorder(region, change_from_2019),
             y = change_from_2019)) +
  geom_col(fill = "#2C5F8A") +
  coord_flip() +
  labs(
    title    = "Change in Household Saving Rate from 2019 to 2020 (EU27)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "Country",
    y        = "Change from 2019 (percentage points)",
    caption  = "Positive values indicate a higher saving rate in 2020 compared with 2019."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    plot.caption  = element_text(color = "grey50", size = 8, hjust = 0)
  )
```

```{r inflation_saving_scatter}
combined %>%
  filter(!is.na(inflation_rate)) %>%
  ggplot(aes(x = inflation_rate, y = saving_rate, color = covid_phase)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(
    title    = "Inflation vs Household Saving Rate (EU27, 2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki and prc_hicp_ainr)",
    x        = "Inflation rate (%)",
    y        = "Gross saving rate (%)",
    color    = "Period",
    caption  = "This visualisation shows correlation only, not causality."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(color = "grey40", size = 10),
    plot.caption    = element_text(color = "grey50", size = 8, hjust = 0),
    legend.position = "bottom"
  )
```

```{r correlation_check}
combined %>%
  filter(!is.na(inflation_rate)) %>%
  group_by(covid_phase) %>%
  summarise(
    correlation = round(cor(inflation_rate, saving_rate), 2),
    n           = n(),
    .groups = "drop"
  )
```

#write here about event analysis: saving rates jumped sharply in 2020 during the COVID shock, remained elevated in early recovery, then declined as the 2022 inflation surge eroded disposable income; use the correlation table output above to state whether the relationship between inflation and saving is positive or negative per phase, and note this is correlation not causation xyz ball


## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: https://github.com/OmarZakk/nieuw_PFE

## 5.2 Reference list

Use APA referencing throughout your document.
