---
title: "Template"
author: "Studentnames and studentnumbers here"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
csl: apa.csl

---
penis cancer test test
# Set-up your environment
شه زود يا ابا الفاضل

```{r package_install, include=FALSE}
install.packages("tidyverse")
install.packages("yaml")
install.packages("rmarkdown")

```

```{r packages}
require(tidyverse)
require(rmarkdown)
require(yaml)
library(tidyr)
library(writexl)
library(readxl)
```

# Title Page
A new sentence
Include your names

Include the tutorial group number

Include your tutorial lecturer's name

# Part 1 - Identify a Social Problem

Use APA referencing throughout your document. [Here's a link to some explanation.](https://www.mendeley.com/guides/apa-citation-guide/)

## 1.1 Describe the Social Problem

Include the following:

-   Why is this relevant? [@smith2020]

-   ...

# Part 2 - Data Sourcing
## 2.1 Load in the data
Preferably from a URL, but if not, make sure to download the data and store it in a shared location that you can load the data in from. Do not store the data in a folder you include in the Github repository!
```{r loading_data}
library(tidyverse)
library(readxl)

# Each teammate sets their own path to the shared download location (NOT in the repo)
data_path <- "C:/Users/Gebruiker/Downloads/"

savings_raw   <- read_xlsx(file.path(data_path, "nasa_10_ki$defaultview_spreadsheet.xlsx"), skip = 9)
inflation_raw <- read_xlsx(file.path(data_path, "prc_hicp_ainr$defaultview_spreadsheet.xlsx"), skip = 9)
```
## 2.2 Provide a short summary of the dataset(s)
```{r summary}
head(savings_raw)
head(inflation_raw)
```
#write here about the metadata: source (Eurostat nasa_10_ki for saving rate, prc_hicp_ainr for inflation), units (saving rate = % of gross disposable income; inflation = HICP annual average index 2025=100, from which we compute annual % change), time coverage 2016-2025, unit of observation = country-year xyz
## 2.3 Describe the type of variables included
Think of things like:
-   Do the variables contain health information or SES information?
-   Have they been measured by interviewing individuals or is the data coming from administrative sources?
#write here about variable types: both are macroeconomic / SES indicators from administrative national accounts data (ESA 2010), not survey/interview data; saving rate and inflation are ratio-level numeric variables; country is a nominal factor; development status is a derived ordinal/categorical variable xyz
# Part 3 - Quantifying
## 3.1 Data cleaning
Say we want to include only larger distances (above 2) in our dataset, we can filter for this.
```{r data_cleaning_build}
year <- c(2018, 2019, 2020, 2021, 2022, 2023, 2024)

savings <- data.frame(
  year        = year,
  Germany     = c(18.24, 17.89, 23.34, 21.96, 18.82, 19.10, 19.93),
  Netherlands = c(13.90, 16.21, 22.41, 19.39, 14.76, 15.01, 16.83),
  Poland      = c( 6.01,  7.18, 13.59,  6.22,  2.16,  6.01,  9.04),
  Romania     = c(-11.06, -2.08,  1.97, -1.69, -7.58, -2.58,  0.05)
)

inflation <- data.frame(
  year        = year,
  Germany     = c(1.91, 1.38, 0.36, 3.20,  8.67,  6.02, 2.49),
  Netherlands = c(1.59, 2.68, 1.12, 2.82, 11.64,  4.09, 3.22),
  Poland      = c(1.17, 2.16, 3.64, 5.23, 13.20, 10.84, 3.66),
  Romania     = c(4.08, 3.91, 2.32, 4.11, 12.02,  9.75, 5.84)
)
```
Please use a separate 'R block' of code for each type of cleaning. So, e.g. one for missing values, a new one for removing unnecessary variables etc.
```{r data_cleaning_reshape}
savings_long   <- pivot_longer(savings,   -year, names_to = "country", values_to = "saving_rate")
inflation_long <- pivot_longer(inflation, -year, names_to = "country", values_to = "inflation_rate")

combined <- left_join(savings_long, inflation_long, by = c("year", "country"))
```
```{r data_cleaning_missing}
sum(is.na(combined))
combined <- combined %>% drop_na()
```
## 3.2 Generate necessary variables
Variable 1
```{r gen_var1}
combined <- combined %>%
  mutate(development = ifelse(country %in% c("Germany", "Netherlands"),
                              "Developed (Western EU)",
                              "Developing (Eastern EU)"))
```
Variable 2
```{r gen_var2}
combined <- combined %>%
  mutate(period = case_when(
    year <  2020 ~ "Pre-COVID",
    year == 2020 ~ "COVID shock",
    year >  2020 ~ "Post-COVID"
  ))

combined$country     <- factor(combined$country,
                               levels = c("Germany", "Netherlands", "Poland", "Romania"))
combined$development <- factor(combined$development)
combined$period      <- factor(combined$period,
                               levels = c("Pre-COVID", "COVID shock", "Post-COVID"))
```
## 3.3 Visualize temporal variation
```{r temporal_plot}
country_colors <- c(
  "Germany"     = "#2C5F8A",
  "Netherlands" = "#4AA8D8",
  "Poland"      = "#E07B39",
  "Romania"     = "#C0392B"
)

ggplot(combined, aes(x = year, y = saving_rate, color = country, linetype = development)) +
  annotate("rect", xmin = 2019.5, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.08) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red", linewidth = 0.7) +
  annotate("text", x = 2020.1, y = -10, label = "COVID-19 (2020)",
           color = "red", size = 3.2, hjust = 0) +
  geom_hline(yintercept = 0, color = "grey40", linewidth = 0.4, linetype = "dotted") +
  geom_line(linewidth = 1) +
  geom_point(size = 2.5) +
  scale_color_manual(values = country_colors, name = "Country") +
  scale_linetype_manual(values = c("Developed (Western EU)" = "solid",
                                   "Developing (Eastern EU)" = "dashed"),
                        name = "Development status") +
  labs(
    title    = "Gross Household Saving Rate by Country (2018-2024)",
    subtitle = "Developed vs. developing EU economies | Source: Eurostat (nasa_10_ki)",
    x        = "Year",
    y        = "Gross saving rate (% of disposable income)",
    caption  = "Note: Provisional data for 2023-2024. Dashed line = COVID-19 shock year."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    plot.caption  = element_text(color = "grey50", size = 8, hjust = 0),
    legend.position = "bottom"
  ) +
  scale_x_continuous(breaks = 2018:2024)
```
#write here about why the temporal plot matters: COVID savings spike visible in all four countries in 2020, sharper in developed economies, then a sharp drop in 2022 as inflation eroded disposable income xyz
## 3.4 Visualize spatial variation
```{r visualise_map}
library(maps)

eu_map <- map_data("world", region = c("Germany", "Netherlands", "Poland", "Romania"))

savings_2024 <- combined %>%
  filter(year == 2024) %>%
  mutate(region = as.character(country))

map_data_joined <- left_join(eu_map, savings_2024, by = "region")

ggplot(map_data_joined, aes(x = long, y = lat, group = group, fill = saving_rate)) +
  geom_polygon(color = "white", linewidth = 0.3) +
  coord_map() +
  scale_fill_gradient2(low = "#C0392B", mid = "#F4E04D", high = "#2C5F8A",
                       midpoint = 10, name = "Saving rate (%)") +
  labs(
    title    = "Gross Household Saving Rate by Country (2024)",
    subtitle = "Spatial comparison across selected EU economies | Source: Eurostat (nasa_10_ki)",
    x        = NULL, y = NULL
  ) +
  theme_void(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    legend.position = "right"
  )
```
#write here about why the spatial plot matters: clear west-east gradient in 2024 saving rates, developed western economies (Germany, Netherlands) saving far more than developing eastern ones (Poland, Romania) xyz
## 3.5 Visualize sub-population variation
What is the saving rate by development status?
```{r visualise_subpopulations}
ggplot(combined, aes(x = development, y = saving_rate, fill = development)) +
  geom_boxplot(alpha = 0.7) +
  geom_jitter(aes(color = country), width = 0.15, size = 2) +
  scale_fill_manual(values = c("Developed (Western EU)" = "#2C5F8A",
                               "Developing (Eastern EU)" = "#E07B39"),
                    guide = "none") +
  scale_color_manual(values = country_colors, name = "Country") +
  labs(
    title = "Distribution of Saving Rates by Development Status (2018-2024)",
    x     = "Development category",
    y     = "Gross saving rate (%)"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold", size = 13),
    legend.position = "right"
  )
```
#write here about why the sub-population plot matters: developed economies cluster at consistently high, positive saving rates; developing economies spread much lower and dip into negative territory (dissaving) xyz
## 3.6 Event analysis
Analyze the relationship between two variables.
```{r analysis}
scale_factor <- 2

strip_labels <- c(
  "Germany"     = "Germany - Developed (Western EU)",
  "Netherlands" = "Netherlands - Developed (Western EU)",
  "Poland"      = "Poland - Developing (Eastern EU)",
  "Romania"     = "Romania - Developing (Eastern EU)"
)

ggplot(combined, aes(x = year)) +
  annotate("rect", xmin = 2019.5, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.07) +
  geom_vline(xintercept = 2020, linetype = "dashed", color = "red", linewidth = 0.6) +
  geom_hline(yintercept = 0, color = "grey60", linewidth = 0.3, linetype = "dotted") +
  geom_line(aes(y = saving_rate, color = country), linewidth = 1.1) +
  geom_point(aes(y = saving_rate, color = country), size = 2.5) +
  geom_line(aes(y = inflation_rate * scale_factor, color = country),
            linewidth = 1.1, linetype = "dotted") +
  geom_point(aes(y = inflation_rate * scale_factor, color = country),
             size = 2.5, shape = 17) +
  scale_y_continuous(
    name = "Gross saving rate (% of disposable income)",
    sec.axis = sec_axis(~ . / scale_factor,
                        name = "HICP inflation rate (% annual change)")
  ) +
  scale_color_manual(values = country_colors, guide = "none") +
  scale_x_continuous(breaks = 2018:2024) +
  facet_wrap(~ country, ncol = 2, labeller = labeller(country = strip_labels)) +
  labs(
    title    = "Household Saving Rate and Inflation by Country (2018-2024)",
    subtitle = "Solid line = saving rate (left axis) | Dotted line, triangle = HICP inflation (right axis)",
    x        = "Year",
    caption  = "Sources: Eurostat nasa_10_ki (saving rate); Eurostat prc_hicp_ainr (HICP inflation).\nDashed red line = COVID-19 shock (2020). Provisional data for 2023-2024."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title         = element_text(face = "bold", size = 13),
    plot.subtitle      = element_text(color = "grey40", size = 9),
    plot.caption       = element_text(color = "grey50", size = 8, hjust = 0),
    axis.title.y.left  = element_text(color = "black", size = 10),
    axis.title.y.right = element_text(color = "grey40", size = 10),
    strip.background   = element_rect(fill = "grey95", color = "grey80"),
    strip.text         = element_text(face = "bold", size = 10),
    panel.spacing      = unit(1.2, "lines")
  )
```
#write here about why the event analysis matters: the inverse relationship between the 2022 inflation spike and the collapse in saving rates, especially severe in developing economies (Poland, Romania) where inflation hit hardest and savings turned negative xyz holy yap



# Part 4 - Discussion

## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: ...

## 5.2 Reference list

Use APA referencing throughout your document.
