---
title: "Template"
author: "Studentnames and studentnumbers here"
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
csl: apa.csl

---

# Set-up your environment

```{r package_install, include=FALSE}
install.packages("tidyverse")
install.packages("yaml")
install.packages("rmarkdown")

```

```{r packages}
require(tidyverse)
require(rmarkdown)
require(yaml)
library(tidyr)
library(writexl)
library(readxl)
```

# Title Page
A new sentence
Include your names

Include the tutorial group number

Include your tutorial lecturer's name

# Part 1 - Identify a Social Problem

Use APA referencing throughout your document. [Here's a link to some explanation.](https://www.mendeley.com/guides/apa-citation-guide/)

## 1.1 Describe the Social Problem

Include the following:

-   Why is this relevant? [@smith2020]

-   ...
# Part 2 - Data Sourcing

## 2.1 Load in the data

```{r loading_data}
library(tidyverse)
library(maps)

savings_clean <- data.frame(
  region = c("Belgium","Bulgaria","Czech Republic","Denmark","Germany",
             "Estonia","Ireland","Greece","Spain","France","Croatia",
             "Italy","Cyprus","Latvia","Lithuania","Luxembourg","Hungary",
             "Malta","Netherlands","Austria","Poland","Portugal","Romania",
             "Slovenia","Slovakia","Finland","Sweden","Norway","Switzerland","Serbia"),
  `2019` = c(12.26, 0.23, 11.85, 7.89, 17.89, 10.32, 13.66, -1.32, 8.66,
             14.62, 9.25, 10.52, 6.47, 8.65, 1.61, 15.48, 16.09, 13.26,
             16.21, 13.23, 7.18, 6.89, -2.08, 13.83, 9.95, 10.60, 17.03,
             13.44, 22.11, 2.88),
  `2020` = c(19.25, -0.49, 18.83, 8.98, 23.34, 13.22, 26.15, 1.76, 17.77,
             20.43, 13.82, 17.86, 13.80, 13.94, 11.70, 26.92, 16.08, 24.69,
             22.41, 19.32, 13.59, 11.84, 1.97, 22.88, 11.71, 14.05, 17.61,
             19.46, 26.60, 9.68),
  `2021` = c(16.80, -5.46, 19.78, 6.43, 21.96, 8.51, 21.63, 5.61, 14.29,
             19.07, 12.74, 15.87, 15.10, 10.46, 10.14, 18.74, 18.49, 23.15,
             19.39, 17.23, 6.22, 10.62, -1.69, 16.98, 11.32, 12.06, 17.53,
             19.35, 26.09, 9.64),
  `2022` = c(13.01, -6.40, 18.47, 10.91, 18.82, 2.00, 14.62, -3.93, 8.95,
             16.86, 7.81, 11.44, 5.55, 3.63, 4.48, 13.48, 15.97, 19.47,
             14.76, 15.27, 2.16, 6.86, -7.58, 14.19, 5.92, 9.88, 15.75,
             11.41, 22.73, 6.36),
  check.names = FALSE
)

savings_long <- savings_clean %>%
  pivot_longer(-region, names_to = "year", values_to = "saving_rate") %>%
  mutate(
    year   = as.integer(year),
    period = case_when(
      year == 2019 ~ "Pre-COVID (2019)",
      year == 2020 ~ "COVID shock (2020)",
      year == 2021 ~ "Post-COVID (2021)",
      year == 2022 ~ "Post-COVID (2022)"
    ),
    period = factor(period, levels = c(
      "Pre-COVID (2019)",
      "COVID shock (2020)",
      "Post-COVID (2021)",
      "Post-COVID (2022)"
    ))
  )
```

## 2.2 Provide a short summary of the dataset(s)

```{r summary}
head(savings_long)
```

#write here about metadata: source is Eurostat (nasa_10_ki), indicator is gross household saving rate as % of disposable income, time coverage 2019-2022, unit of observation is country-year xyz

## 2.3 Describe the type of variables included

#write here: saving rate is a numeric ratio-level variable, country is nominal, year is interval; data comes from administrative national accounts (ESA 2010), not surveys xyz

---

# Part 3 - Quantifying

## 3.1 Data cleaning

```{r data_cleaning}
savings_long <- savings_long %>%
  drop_na(saving_rate)
```

## 3.2 Generate necessary variables

```{r gen_var1}
savings_long <- savings_long %>%
  mutate(development = ifelse(
    region %in% c("Germany", "Netherlands", "France", "Belgium",
                  "Austria", "Sweden", "Denmark", "Finland",
                  "Ireland", "Luxembourg"),
    "Developed (Western EU)",
    "Developing (Eastern EU)"
  ))
```

## 3.3 Visualize temporal variation

#write here about temporal variation xyz

## 3.4 Visualize spatial variation

```{r visualise_map, fig.width=14, fig.height=10}
europe_map <- map_data("world")

map_joined <- left_join(europe_map, savings_long, by = "region") %>%
  filter(!is.na(period))

ggplot(map_joined, aes(x = long, y = lat, group = group, fill = saving_rate)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(ratio = 1.3, xlim = c(-25, 45), ylim = c(34, 72)) +
  scale_fill_gradient2(
    low      = "darkred",
    mid      = "yellow",
    high     = "darkgreen",
    midpoint = 10,
    na.value = "grey90",
    name     = "Saving rate (%)"
  ) +
  facet_wrap(~ period, ncol = 2, drop = TRUE) +
  labs(
    title    = "Gross Household Saving Rate across Europe (2019-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki) | Grey = no data available",
    x = NULL, y = NULL
  ) +
  theme_void() +
  theme(
    plot.title      = element_text(face = "bold", size = 14),
    plot.subtitle   = element_text(color = "grey40", size = 11),
    strip.text      = element_text(face = "bold", size = 12),
    legend.position = "bottom",
    legend.key.width = unit(2, "cm")
  )
```
#write here about spatial variation: in 2020 the whole map turns greener as lockdowns forced saving across all countries; by 2022 eastern countries turn red as inflation eroded household income xyz

## 3.5 Visualize sub-population variation

#write here about sub-population variation xyz

## 3.6 Event analysis

#write here about event analysis xyz

## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: ...

## 5.2 Reference list

Use APA referencing throughout your document.
