---
title: "Template"
<<<<<<< HEAD
<<<<<<< HEAD
<<<<<<< HEAD
author: "Omar Zakkouralashy (2889420), Anna Mishmish-Olaru (2881939), Abhigyan Hazarika (student number), Andrea Tzima or Xhima,Raluca Naumcef(2888853), kyle"
=======
<<<<<<< HEAD
author: "Omar Zakkouralashy (2889420), Anna Olaru (2881939), Abhigyan Hazarika (student number), Andrea Tzima or Xhima,Raluca Naumcef, kyle"
=======
author: "Omar Zakkouralashy (2889420), Anna Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima,Raluca Naumcef, Kyle van Emmerik (2897218)"
>>>>>>> b62ff297222185e7bc4e016fd486e3fee4b2a663
=======
=======
author: "Omar Zakkouralashy (2889420), Anna Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima,Raluca Naumcef, Kyle van Emmerik (2897218)"
=======
>>>>>>> b62ff297222185e7bc4e016fd486e3fee4b2a663
author: "Omar Zakkouralashy (2889420), Anna Mishmish-Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima (2885530),Raluca Naumcef, Kyle van Emmerik (2897218)"
>>>>>>> d0097343dd07037992a3f4c54a811f9a9756b238
>>>>>>> a66092d3728bb9fa977f156a6f3c809c2eb7c25b
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
csl: apa.csl

---

# Set-up your environment

```{r package_install, include=FALSE}
install.packages("tidyverse")
install.packages("yaml")
install.packages("rmarkdown")
install.packages("maps")
install.packages("eurostat")
```

```{r packages}
require(tidyverse)
require(rmarkdown)
require(yaml)
library(tidyr)
library(writexl)
library(readxl)
require()
```

# Title Page
The effect of COVID on household savings in the European Union - A social problem.

Omar, Anna, Abhigyan, Raluca, Andrea, Kyle

TG 5 - G 1

Tutorial teacher: Simon 

Simon Loop
Simon Loop

# Part 1 - Identify a Social Problem
Aside from being a world wide health crisis, the Covid-19 pandemic, simultaneously, limited individual spending and generated a sense of financial insecurity among all classes. Through a survey on individuals in eight countries across the world, Deloitte concluded that, while prior to the pandemic a large portion of income surplus went into consumption for most participants, attitudes have now shifted towards dedicating that surplus to savings and investments.
The restrictions imposed by governments during lockdown limited spending opportunities, such as traveling and other non-essentials, which paired with government subsidies in many countries (Deloitte, 2022), increased savings. Additionally, the pandemic brought a sense of economic uncertainty regarding employment and future economic and health stability, making individuals crave solidity, safety, and reassurance, reflected onto their growing savings (Deloitte, 2022). 
We see this represented quite well in countries like the Netherlands and Germany, experiencing two of the highest household savings even after the lockdown, 24% and 23.4% respectively (Eurostat), because of great government support causing consumption to fall faster than income. 
On the other hand, countries like Romania faced a decrease in savings after the pandemic, due to fewer financial resources and rising living costs, also reflected in the country’s economic status . In these countries savings were predominantly used to cope with the post-Covid inflation period (Belbe et al, 2024)
Thus, observing household savings prior, during, and after the pandemic, provides insights into the issue of financial vulnerability as a result of the pandemic, as well as inequality between European countries. The issue remains relevant today, because the change in consumption mindsets, due to Covid, seems persistent even well beyond the recovery phase and into present times. 


# Part 2 - Data Sourcing

## 2.1 Load in the data

```{r loading_data}
library(maps)
library(eurostat)
library(grid)

dat_savings   <- get_eurostat("nasa_10_ki",    time_format = "num", cache = TRUE)
dat_inflation <- get_eurostat("prc_hicp_ainr", time_format = "num", cache = TRUE)
```

## 2.2 Provide a short summary of the dataset(s)

```{r summary}
dataset_summary <- data.frame(
  Dataset    = c("nasa_10_ki", "prc_hicp_ainr"),
  Source     = c("Eurostat", "Eurostat"),
  Indicator  = c("Gross household saving rate",
                 "HICP inflation rate"),
  Unit       = c("% of gross disposable income",
                 "Annual average % change"),
  Level      = c("Country-year", "Country-year"),
  Years_used = c("2016-2022", "2016-2022"),
  Limitation = c("Country-level only, no household-level detail",
                 "National average, not household-specific")
)

dataset_summary
```

Two datasets are used in this analysis, both sourced from Eurostat (the statistical office of the European Union) and downloaded in June 2026.

The HICP – Inflation Rate dataset (code: tec00118, last updated 20/05/2026) has been computed using the Eurostat dataset prc_hicp_ainr. It reports the annual average rate of change in consumer prices across all HICP categories, covering EU averages and individual EU member states. The HICP is a standardized measure of consumer price inflation in Europe ensuring consistency and comparability of data across participating countries. The data runs annually from 2014 to 2025, however for the sake of our research we use data from 2016 to 2022.

The household saving rate dataset (code: tec00131, last updated 04/06/2026) is compiled from annual national accounts data (source: nasa_10_ki) and covers individual EU participating countries as well as EU and euro area aggregates, running annually from 2014 to 2025 however again the data used is from 2016 to 2022. It measures the share of gross disposable income that households set aside rather than spend, expressed as a percentage. The sector includes both households and non-profit institutions serving households.


## 2.3 Describe the type of variables included

The key numeric variables are the saving rate and inflation rate. Both are numeric variables expressed as percentages, allowing direct comparison across countries and time periods

Country (GEO) is a nominal variable, identifying each EU participating country and averages across Europe. Time is annual in both datasets, ranging from 2014 to 2025, and is treated as an interval variable where the distance between years is equal and meaningful. Later variables created in 3.2 are categorical. 

Both variables come from administrative and index-based sources rather than surveys, which limits typical survey sampling error but means values depend on the quality and timeliness of national accounts reporting by each member state.


---

# Part 3 - Quantifying

## 3.1 Data cleaning

```{r data_cleaning_savings}
eu27 <- c(
  "AT", "BE", "BG", "HR", "CY", "CZ", "DK", "EE", "FI", "FR",
  "DE", "EL", "HU", "IE", "IT", "LV", "LT", "LU", "MT", "NL",
  "PL", "PT", "RO", "SK", "SI", "ES", "SE"
)

years_used <- 2016:2022

savings_clean <- dat_savings %>%
  filter(
    na_item == "SRG_S14_S15",
    sector  == "S14_S15",
    unit    == "PC",
    geo     %in% eu27
  ) %>%
  mutate(year = as.integer(TIME_PERIOD)) %>%
  filter(year %in% years_used) %>%
  select(geo, year, saving_rate = values) %>%
  drop_na(saving_rate)
```

```{r data_cleaning_inflation}
inflation_clean <- dat_inflation %>%
  filter(
    coicop18 == "AP",
    unit     == "RCH_A_AVG",
    geo      %in% eu27
  ) %>%
  mutate(year = as.integer(TIME_PERIOD)) %>%
  filter(year %in% years_used) %>%
  select(geo, year, inflation_rate = values) %>%
  drop_na(inflation_rate)
```

```{r data_cleaning_merge}
combined <- left_join(savings_clean, inflation_clean, by = c("geo", "year"))
```

```{r data_cleaning_check}
combined %>%
  summarise(
    countries         = n_distinct(geo),
    first_year        = min(year),
    last_year         = max(year),
    observations      = n(),
    missing_saving    = sum(is.na(saving_rate)),
    missing_inflation = sum(is.na(inflation_rate))
  )
```

## 3.2 Generate necessary variables

```{r gen_var1}
country_names <- c(
  AT = "Austria",     BE = "Belgium",      BG = "Bulgaria",
  HR = "Croatia",     CY = "Cyprus",       CZ = "Czech Republic",
  DK = "Denmark",     EE = "Estonia",      FI = "Finland",
  FR = "France",      DE = "Germany",      EL = "Greece",
  HU = "Hungary",     IE = "Ireland",      IT = "Italy",
  LV = "Latvia",      LT = "Lithuania",    LU = "Luxembourg",
  MT = "Malta",       NL = "Netherlands",  PL = "Poland",
  PT = "Portugal",    RO = "Romania",      SK = "Slovakia",
  SI = "Slovenia",    ES = "Spain",        SE = "Sweden"
)

combined <- combined %>%
  mutate(
    region = country_names[as.character(geo)],
    region_group = case_when(
      geo %in% c("AT", "BE", "DE", "FR", "IE", "LU", "NL") ~ "Western Europe",
      geo %in% c("DK", "FI", "SE")                          ~ "Northern Europe",
      geo %in% c("CY", "EL", "ES", "IT", "MT", "PT")        ~ "Southern Europe",
      geo %in% c("BG", "CZ", "EE", "HR", "HU", "LV",
                 "LT", "PL", "RO", "SI", "SK")               ~ "Central/Eastern Europe"
    ),
    covid_phase = case_when(
      year <= 2019 ~ "Pre-COVID",
      year == 2020 ~ "COVID shock",
      year >= 2021 ~ "Recovery / inflation period"
    ),
    covid_phase = factor(covid_phase, levels = c(
      "Pre-COVID",
      "COVID shock",
      "Recovery / inflation period"
    )),
    map_period = case_when(
      year == 2019 ~ "Pre-COVID (2019)",
      year == 2020 ~ "COVID shock (2020)",
      year == 2021 ~ "Post-COVID (2021)",
      year == 2022 ~ "Post-COVID (2022)"
    ),
    map_period = factor(map_period, levels = c(
      "Pre-COVID (2019)",
      "COVID shock (2020)",
      "Post-COVID (2021)",
      "Post-COVID (2022)"
    ))
  ) %>%
  drop_na(region, region_group)
```

```{r gen_var2}
baseline_2019 <- combined %>%
  filter(year == 2019) %>%
  select(geo, saving_rate_2019 = saving_rate)

combined <- combined %>%
  left_join(baseline_2019, by = "geo") %>%
  mutate(change_from_2019 = saving_rate - saving_rate_2019)

savings_clean <- combined
```

## 3.3 Visualize temporal variation

```{r temporal_plot}
eu_mean <- combined %>%
  group_by(year) %>%
  summarise(
    mean_saving_rate    = mean(saving_rate,    na.rm = TRUE),
    mean_inflation_rate = mean(inflation_rate, na.rm = TRUE),
    n_countries         = n_distinct(geo),
    .groups = "drop"
  )

ggplot(eu_mean, aes(x = year, y = mean_saving_rate)) +
  geom_line(color = "#2C5F8A", linewidth = 1) +
  geom_point(color = "#2C5F8A", size = 3) +
  annotate("rect", xmin = 2019.5, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.08) +
  geom_vline(xintercept = 2020, linetype = "dashed",
             color = "red", linewidth = 0.7) +
  annotate("text", x = 2020.1, y = 10,
           label = "COVID-19 (2020)", color = "red",
           size = 3.2, hjust = 0) +
  labs(
    title    = "EU27 Mean Household Saving Rate (2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "Year",
    y        = "Mean saving rate (%)",
    caption  = "Note: Unweighted mean across EU27 countries with available data."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    plot.caption  = element_text(color = "grey50", size = 8, hjust = 0)
  ) +
  scale_x_continuous(breaks = 2016:2022)
```

#The saving behaviour of households has changed dramatically. In 2020, the average saving rate of the EU27 shot up as the COVID‑19 lockdowns limited consumption possibilities and increased involuntary savings. Restrictions were lifted in 2021 and the saving rate has gradually declined as spending habits have normalised. Then high inflation ate into real disposable incomes, and households accumulated less savings to support consumption.

## 3.4 Visualize spatial variation

```{r visualise_map, fig.width=14, fig.height=10}
europe_map <- map_data("world")

unmatched <- combined %>%
  distinct(region) %>%
  anti_join(europe_map %>% distinct(region), by = "region")
unmatched

map_joined <- left_join(europe_map, combined, by = "region") %>%
  filter(!is.na(map_period))

ggplot(map_joined, aes(x = long, y = lat, group = group, fill = saving_rate)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(ratio = 1.3, xlim = c(-25, 45), ylim = c(34, 72)) +
  scale_fill_gradient2(
    low      = "darkred",
    mid      = "yellow",
    high     = "darkgreen",
    midpoint = median(combined$saving_rate, na.rm = TRUE),
    na.value = "grey90",
    name     = "Saving rate (%)"
  ) +
  facet_wrap(~ map_period, ncol = 2, drop = TRUE) +
  labs(
    title    = "Gross Household Saving Rate across EU27 (2019-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki) | Grey = no data available",
    x = NULL, y = NULL,
    caption  = "Saving rate = gross household saving as % of disposable income. Midpoint set at EU27 median."
  ) +
  theme_void() +
  theme(
    plot.title       = element_text(face = "bold", size = 14),
    plot.subtitle    = element_text(color = "grey40", size = 11),
    plot.caption     = element_text(color = "grey50", size = 8, hjust = 0),
    strip.text       = element_text(face = "bold", size = 12),
    legend.position  = "bottom",
    legend.key.width = grid::unit(2, "cm")
  )
```

#The spatial variation is also quite telling: For most of the EU27 countries, the map is a lot greener in 2020, reflecting the universal rise in saving as lockdowns curtailed consumption everywhere. By 2022 a clear east-west divide appears, with the central and eastern member states turning red. This reflects steeper declines in saving rates as high inflation more sharply eroded household purchasing power in those economies.

## 3.5 Visualize sub-population variation

```{r visualise_subpopulations}
combined %>%
  ggplot(aes(x = covid_phase, y = saving_rate, fill = region_group)) +
  geom_boxplot(alpha = 0.7) +
  labs(
    title    = "Saving Rate by Region Group and COVID Phase (EU27, 2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "COVID phase",
    y        = "Gross saving rate (%)",
    fill     = "Region group",
    caption  = "Groups are broad geographical categories and do not imply causal explanations."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(color = "grey40", size = 10),
    plot.caption    = element_text(color = "grey50", size = 8, hjust = 0),
    axis.text.x     = element_text(angle = 20, hjust = 1),
    legend.position = "bottom"
  )
```

#The difference in sub-populations is a structural divide that persists over time with higher saving rates persistently higher in Western and northern Europe and much lower saving rates in southern and central/eastern Europe. The COVID shock of 2020 temporarily closed some of these gaps, with sharp increases in saving rates across all groups while consumption fell sharply. But the post-COVID recovery has been uneven and pre-existing regional differences have re-emerged as countries have been affected by inflation and income pressures to different degrees.

## 3.6 Event analysis

```{r analysis}
combined %>%
  ggplot(aes(x = covid_phase, y = saving_rate, fill = covid_phase)) +
  geom_boxplot(alpha = 0.7) +
  scale_fill_manual(
    values = c(
      "Pre-COVID"                   = "#2C5F8A",
      "COVID shock"                 = "#C0392B",
      "Recovery / inflation period" = "#E07B39"
    ),
    guide = "none"
  ) +
  labs(
    title    = "Household Saving Rate around the COVID-19 Shock (EU27, 2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "Period",
    y        = "Gross saving rate (%)",
    caption  = "This plot shows association only; it does not prove COVID caused the observed changes."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    plot.caption  = element_text(color = "grey50", size = 8, hjust = 0)
  )
```

```{r change_from_2019_plot}
combined %>%
  filter(year == 2020) %>%
  ggplot(aes(x = reorder(region, change_from_2019),
             y = change_from_2019)) +
  geom_col(fill = "#2C5F8A") +
  coord_flip() +
  labs(
    title    = "Change in Household Saving Rate from 2019 to 2020 (EU27)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "Country",
    y        = "Change from 2019 (percentage points)",
    caption  = "Positive values indicate a higher saving rate in 2020 compared with 2019."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10),
    plot.caption  = element_text(color = "grey50", size = 8, hjust = 0)
  )
```

```{r inflation_saving_scatter}
combined %>%
  filter(!is.na(inflation_rate)) %>%
  ggplot(aes(x = inflation_rate, y = saving_rate, color = covid_phase)) +
  geom_point(alpha = 0.7, size = 2) +
  labs(
    title    = "Inflation vs Household Saving Rate (EU27, 2016-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki and prc_hicp_ainr)",
    x        = "Inflation rate (%)",
    y        = "Gross saving rate (%)",
    color    = "Period",
    caption  = "This visualisation shows correlation only, not causality."
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title      = element_text(face = "bold", size = 13),
    plot.subtitle   = element_text(color = "grey40", size = 10),
    plot.caption    = element_text(color = "grey50", size = 8, hjust = 0),
    legend.position = "bottom"
  )
```

```{r correlation_check}
combined %>%
  filter(!is.na(inflation_rate)) %>%
  group_by(covid_phase) %>%
  summarise(
    correlation = round(cor(inflation_rate, saving_rate), 2),
    n           = n(),
    .groups = "drop"
  )
```

#It can be observed from event analysis that there exists a definite three phase dynamics in the relationship between inflation and saving. The first phase is represented by the occurrence of the COVID shock in 2020 where consumption became difficult for households leading to a situation where people were forced to save. This led to an increase in saving rates while in the second phase, savings declined along with real disposable income due to increased inflation in 2022. Here is how it appears in the correlation table. In the COVID shock phase, there is a positive correlation between inflation and saving which implies that both variables are positively correlated in the given phase. In the inflation phase, the relationship between inflation and saving is negative which implies that saving is negatively correlated with inflation.


## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: https://github.com/OmarZakk/nieuw_PFE

## 5.2 Reference list

Use APA referencing throughout your document.
