---
title: "Template"
<<<<<<< HEAD
author: "Omar Zakkouralashy (2889420), Anna Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima,Raluca Naumcef, kyle"
=======
author: "Omar Zakkouralashy (2889420), Anna Mishmish-Olaru (2881939), Abhigyan Hazarika (2888419), Andrea Tzima or Xhima (2885530),Raluca Naumcef, kyle"
>>>>>>> d0097343dd07037992a3f4c54a811f9a9756b238
date: "`r Sys.Date()`"
output: pdf_document
bibliography: references.bib
csl: apa.csl

---

# Set-up your environment

```{r package_install, include=FALSE}
install.packages("tidyverse")
install.packages("yaml")
install.packages("rmarkdown")

```

```{r packages}
require(tidyverse)
require(rmarkdown)
require(yaml)
library(tidyr)
library(writexl)
library(readxl)
```

# Title Page
The effect of COVID on household savings in the European Union - A social problem.

Omar, Anna, Abhigyan, Raluca, Andrea, Kyle

TG 5 - G 1

Tutorial teacher: Simon 

Simon Loop
Simon Loop

# Part 1 - Identify a Social Problem



## 1.1 Describe the Social Problem

Include the following:

-   Why is this relevant? [@smith2020]

-   ...
# Part 2 - Data Sourcing

## 2.1 Load in the data

```{r loading_data}
library(tidyverse)
library(maps)
library(eurostat)

dat <- get_eurostat("nasa_10_ki")
```

## 2.2 Provide a short summary of the dataset(s)

```{r summary}
head(dat)
str(dat)
```

#write here about metadata: source is Eurostat (nasa_10_ki), indicator is gross household saving rate as % of disposable income, time coverage 2016-2025, unit of observation is country-year xyz

## 2.3 Describe the type of variables included

#write here: saving rate is a numeric ratio-level variable, country is nominal, year is interval; data comes from administrative national accounts (ESA 2010), not surveys xyz

---

# Part 3 - Quantifying

## 3.1 Data cleaning

```{r data_cleaning}
savings_clean <- dat %>%
  filter(
    na_item == "SRG_S14",
    sector  == "S14",
    unit    == "PC"
  ) %>%
  mutate(year = as.integer(format(TIME_PERIOD, "%Y"))) %>%
  filter(year %in% c(2019, 2020, 2021, 2022)) %>%
  select(geo, year, saving_rate = values) %>%
  drop_na(saving_rate)
```

## 3.2 Generate necessary variables

```{r gen_var1}
country_names <- c(
  BE = "Belgium",      BG = "Bulgaria",    CZ = "Czech Republic",
  DK = "Denmark",      DE = "Germany",     EE = "Estonia",
  IE = "Ireland",      EL = "Greece",      ES = "Spain",
  FR = "France",       HR = "Croatia",     IT = "Italy",
  CY = "Cyprus",       LV = "Latvia",      LT = "Lithuania",
  LU = "Luxembourg",   HU = "Hungary",     MT = "Malta",
  NL = "Netherlands",  AT = "Austria",     PL = "Poland",
  PT = "Portugal",     RO = "Romania",     SI = "Slovenia",
  SK = "Slovakia",     FI = "Finland",     SE = "Sweden",
  NO = "Norway",       CH = "Switzerland", RS = "Serbia"
)

savings_clean <- savings_clean %>%
  mutate(
    region = country_names[as.character(geo)],
    period = case_when(
      year == 2019 ~ "Pre-COVID (2019)",
      year == 2020 ~ "COVID shock (2020)",
      year == 2021 ~ "Post-COVID (2021)",
      year == 2022 ~ "Post-COVID (2022)"
    ),
    period = factor(period, levels = c(
      "Pre-COVID (2019)",
      "COVID shock (2020)",
      "Post-COVID (2021)",
      "Post-COVID (2022)"
    ))
  ) %>%
  drop_na(region)
```

## 3.3 Visualize temporal variation

```{r temporal_plot}
eu_mean <- dat %>%
  filter(
    na_item == "SRG_S14",
    sector  == "S14",
    unit    == "PC"
  ) %>%
  mutate(year = as.integer(format(TIME_PERIOD, "%Y"))) %>%
  filter(year %in% c(2019, 2020, 2021, 2022)) %>%
  drop_na(values) %>%
  group_by(year) %>%
  summarise(mean_saving_rate = mean(values))

ggplot(eu_mean, aes(x = year, y = mean_saving_rate)) +
  geom_line(color = "#2C5F8A", linewidth = 1) +
  geom_point(color = "#2C5F8A", size = 3) +
  annotate("rect", xmin = 2019.5, xmax = 2020.5,
           ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.08) +
  geom_vline(xintercept = 2020, linetype = "dashed",
             color = "red", linewidth = 0.7) +
  annotate("text", x = 2020.1, y = 10,
           label = "COVID-19 (2020)", color = "red",
           size = 3.2, hjust = 0) +
  labs(
    title    = "EU Mean Household Saving Rate (2019-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki)",
    x        = "Year",
    y        = "Mean saving rate (%)"
  ) +
  theme_classic(base_size = 12) +
  theme(
    plot.title    = element_text(face = "bold", size = 13),
    plot.subtitle = element_text(color = "grey40", size = 10)
  ) +
  scale_x_continuous(breaks = c(2019, 2020, 2021, 2022))
```

#write here about temporal variation: the EU mean saving rate spiked sharply in 2020 as COVID lockdowns suppressed consumption, then gradually declined as economies reopened and inflation eroded disposable income xyz

## 3.4 Visualize spatial variation

```{r visualise_map, fig.width=14, fig.height=10}
europe_map <- map_data("world")

map_joined <- left_join(europe_map, savings_clean, by = "region") %>%
  filter(!is.na(period))

ggplot(map_joined, aes(x = long, y = lat, group = group, fill = saving_rate)) +
  geom_polygon(color = "white", linewidth = 0.2) +
  coord_fixed(ratio = 1.3, xlim = c(-25, 45), ylim = c(34, 72)) +
  scale_fill_gradient2(
    low      = "darkred",
    mid      = "yellow",
    high     = "darkgreen",
    midpoint = 10,
    na.value = "grey90",
    name     = "Saving rate (%)"
  ) +
  facet_wrap(~ period, ncol = 2, drop = TRUE) +
  labs(
    title    = "Gross Household Saving Rate across Europe (2019-2022)",
    subtitle = "Source: Eurostat (nasa_10_ki) | Grey = no data available",
    x = NULL, y = NULL
  ) +
  theme_void() +
  theme(
    plot.title       = element_text(face = "bold", size = 14),
    plot.subtitle    = element_text(color = "grey40", size = 11),
    strip.text       = element_text(face = "bold", size = 12),
    legend.position  = "bottom",
    legend.key.width = unit(2, "cm")
  )
```

#write here about spatial variation: in 2020 the map turns greener across all countries as lockdowns forced saving; by 2022 eastern countries turn red as inflation eroded household income xyz

## 3.5 Visualize sub-population variation

#write here about sub-population variation xyz

## 3.6 Event analysis

#write here about event analysis xyz

#write here about event analysis xyz
```
#write here about spatial variation: in 2020 the whole map turns greener as lockdowns forced saving across all countries; by 2022 eastern countries turn red as inflation eroded household income xyz

## 3.5 Visualize sub-population variation

#write here about sub-population variation xyz

## 3.6 Event analysis

#write here about event analysis xyz

## 4.1 Discuss your findings

# Part 5 - Reproducibility

## 5.1 Github repository link

Provide the link to your PUBLIC repository here: ...

## 5.2 Reference list

Use APA referencing throughout your document.
